---
title: Flow analysis of scCloneSelect (Exp_20210925_1)
subtitle: Plasmid vs lentivirus delivery in mESCs
author: "Soh Ishiguro"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: FALSE
    toc_depth: 3
    toc_collapsed: FALSE
    number_sections: yes
    keep_md: FALSE
    fig_caption: yes
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(
    comment = "#>",
    out.width = "50%",
    dpi = 300,
    warning = FALSE,
    message = FALSE
)
```

# Overview
mESC flow assay was performed by RinaS (Exp_20210719_1) with the following experimental condition: `Exp_20210925_1_metadata.tsv`

Assay was peformed with n=2, to see GFP activation in scCloneSelect construct when we use either plasmid transfection or lenti delivery.

# Analysis
```{r}
library(ggplot2)
library(RColorBrewer)
library(ggcyto)
library(CytoExploreR)
library(flowWorkspace)
library(dplyr)
library(rmarkdown)
library(scales)
library(ggridges)
```

Load samples and create GatingSet object.
```{r}
cs <- load_cytoset_from_fcs(path = "./fcs")
gs <- GatingSet(cs)
sampleNames(gs)
markernames(gs)
cs_get_cytoframe(cs, "Tube.1.fcs") # check data
```
A total of `r length(sampleNames(gs))` samples were loaded.

Apply transformation for fluorescence channels
```{r}
trans.log <- cyto_transformer_log(gs, channels = c("FL1-A", "FL5-A"), plot = FALSE)
gs.trans <- cyto_transform(gs, trans = trans.log, plot = FALSE)
```

Assign metadata to the `gs.trans`
```{r}
meta.df <- read.csv("./Exp_20210925_1_metadata.tsv", header = TRUE, sep = "\t")
sample.df <- cyto_details(gs.trans)
merged.df <- merge(meta.df, sample.df, by.x = "File", by.y = "name")
cyto_details(gs.trans)$ExpID <- merged.df$ExpID
cyto_details(gs.trans)$Cell_line <- merged.df$Cell_line
cyto_details(gs.trans)$gRNA <- merged.df$gRNA
cyto_details(gs.trans)$Rep <- merged.df$Rep
cyto_details(gs.trans)$Sample_type <- merged.df$Sample_type
cyto_details(gs.trans)$Target_type <- merged.df$Target_type
cyto_details(gs.trans)$Delivery_method <- merged.df$Delivery_method
cyto_details(gs.trans)$Cell_type <- merged.df$Cell_type
```

Check all metatada and fcs files assigned.
```{r}
cyto_details(gs.trans) %>%
    arrange(name) %>%
    paged_table()

gs_pop_get_stats(gs.trans) %>% paged_table()
```

Create manual gate
```{r, eval=FALSE}
# Live cell gate
cyto_gate_draw(
    gs.trans,
    parent = "root",
    alias = "P1",
    channels = c("FSC-A", "SSC-A"),
    gatingTemplate = "Exp20210925_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 1 * 10^6),
    type = "polygon"
)
```
```{r, eval=FALSE}
# Single-cell gate
cyto_gate_draw(
    gs.trans,
    parent = "P1",
    alias = "P2",
    channels = c("FSC-A", "FSC-Width"),
    gatingTemplate = "Exp20210925_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 6 * 10^3),
    type = "polygon"
)
```

```{r, eval=FALSE}
# GFP gate
cyto_gate_edit(
    gs.trans["Tube.1.fcs"], # use it for negative control reference
    parent = "P2",
    alias = "GFP",
    channels = c("FL1-A", "FSC-A"),
    gatingTemplate = "Exp20210925_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^7),
    ylim = c(-10, 1 * 10^6),
    type = "interval"
)
```

Apply /P1/P2/GFP gate for each sample 
```{r}
gt <- gatingTemplate("Exp20210925_Gate_03152022.csv")
gates <- flowWorkspace::gs_get_pop_paths(gs)
if ("/P1/P2/GFP" %in% gates) {
    gs_pop_remove(gs.trans, "P1") # remove existing gates
}
gt_gating(gt, gs.trans)
gs_get_pop_paths(gs.trans)
cyto_plot_gating_scheme(gs.trans[[1]])
```

Merge metadata and GFP stat
```{r}
GFP.stat <- gs_pop_get_stats(gs.trans, node = "GFP", type = "percent") %>% mutate(percent = percent * 100)
df.stat <- merge(cyto_details(gs.trans), GFP.stat, by.x = "name", by.y = "sample")
df.stat <- df.stat %>% mutate(col_var = paste(Cell_line, Target_type, sep = ":"))

# export to csv
write.csv(x = df.stat, file = "GFP_activation_03172022.csv", quote = FALSE, row.names = FALSE)
```

