---
title: "Flow data analysis (CA058) - indel reporter assay"
author: "Soh Ishiguro"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: FALSE
    toc_depth: 3
    toc_collapsed: FALSE
    number_sections: yes
    keep_md: yes
    fig_caption: yes
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(
    comment = "#>",
    out.width = "50%",
    dpi = 200
)
```

# Project overview
- Project: CloneSelect, indel reporter flow assay (WT001 and WT001 target sequence)
- Experimental ID: CA058
- Project path: ``r getwd()``

```{r message=FALSE}
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(ggcyto)
library(CytoExploreR)
library(flowWorkspace)
library(flowCore)
library(scales)
```

# Loading data
```{r}
setwd("./notebook/")
fcs.file.path <- "../fcs"
cs <- load_cytoset_from_fcs(path = fcs.file.path)
gs <- GatingSet(cs)
gs.copy <- gs_clone(gs) # original copy
gs_pop_get_stats(gs)
```

# Data transformation
```{r}
trans.log <- cyto_transformer_log(gs, channels = c("FL1-A"), plot = FALSE)
gs.trans <- cyto_transform(gs, trans = trans.log, plot = FALSE)
exprs(gh_pop_get_data(gs.trans, "root")) %>% head()
```

# Add metadata
Add some experimental metadata to the GatingSet object
```{r}
cyto_details(gs.trans)$sample_type <- c(rep("treatment", 12), "control", "control")
cyto_details(gs.trans)$rep <- c(rep(c("rep1", "rep2", "rep3"), 4), NA, NA)
cyto_details(gs.trans)$cell_line <- c(rep("WT001", 6), rep("WT002", 6), NA, NA)
cyto_details(gs.trans)$gRNA <- c(rep(c(rep("gRNA_WT001", 3), rep("gRNA_WT002", 3)), 2), NA, NA)
cyto_details(gs.trans)$gRNA_type <- c(rep("gRNA_OT", 3), rep("gRNA_NT", 3), rep("gRNA_NT", 3), rep("gRNA_OT", 3), NA, NA)
cyto_details(gs.trans)
```

# Manual Gating
```{r, eval=FALSE}
cyto_gate_draw(
    gs.trans,
    parent = "root", alias = "P1",
    channels = c("FSC-A", "SSC-A"),
    gatingTemplate = "CA058_02042022_Gate.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 1 * 10^6),
    type = "polygon"
)
gs_pop_get_count_fast(gs.trans, statistic = "freq")

cyto_gate_draw(
    gs.trans,
    parent = "P1",
    alias = "GFP",
    channels = c("FL1-A", "FSC-A"),
    gatingTemplate = "CA058_02042022_Gate.csv",
    xlim = c(-10, 1 * 10^7),
    ylim = c(-10, 1 * 10^6),
    type = "interval"
)
```

```{r, message=FALSE}
cyto_gatingTemplate_apply(gs.trans, "CA058_02042022_Gate.csv")

# export
# save_gs(gs.trans, path = file.path("./gs_achieve"))

gs_pop_get_stats(gs.trans, type = "percent", nodes = c("GFP"))
gs.treatment <- subset(gs.trans, sample_type == "treatment")
gs.treatment %>%
    cyto_details() %>%
    rmarkdown::paged_table()
```

```{r}
cyto_plot_save(save_as = "plots/gating_scheme_NC.pdf", height = 4, width = 8)
cyto_plot_gating_scheme(gs.trans["NC1.fcs"])
cyto_plot_save(save_as = "plots/gating_scheme_GFP.pdf", height = 4, width = 8)
cyto_plot_gating_scheme(gs.trans["GFP2.fcs"])
```

```{r}
ggcyto(gs.trans["NC1.fcs"], subset = "root", aes(y = "FSC-A", x = "SSC-A")) +
    geom_hex(bins = 128, aes(fill = ..density..)) +
    geom_gate() +
    geom_stats() +
    scale_x_logicle() +
    scale_y_logicle()
```

# Visualization
This is magic number for ggplot2, which is compatible with keynote object size (9pt font and 0.75 line width)
```{r}
FONT.SIZE <- 7
LINE.W <- 0.232 # 0.5pt in Keynote
```

## Scatter plot
Obtain GFP min/max threshold value froom the gatingtemplate
```{r}

inverse.trans_log <- function(x) {
    trans <- flowjo_log_trans()
    inverse.trans <- trans[["inverse"]]
    return(inverse.trans(x))
}

gt <- gatingTemplate("CA058_02042022_Gate.csv") # TODO: access through this object
gfp.min <- 0.85164693891806 # copied from a file
gfp.max <- 1.55871208947202
```

Plot data to check all replicates
```{r, out.width="120%"}
ggcyto(gs.treatment, subset = "P1", aes(y = "FSC-A", x = "FL1-A")) +
    facet_wrap(gRNA ~ cell_line + rep, ncol = 3) +
    geom_hex(bins = 64, aes(fill = ..density..)) +
    geom_gate("GFP") +
    geom_stats(gate = "GFP", digits = 2) +
    scale_fill_gradientn(colours = brewer.pal(n = 4, name = "RdYlBu")) +
    axis_x_inverse_trans() +
    axis_y_inverse_trans() +
    coord_cartesian(ylim = c(-10, 600000), xlim = c(0, 1.4)) +
    theme_classic() +
    theme(strip.background = element_blank())
```

The following is for Supplemenetary Figure 8
```{r, out.width="60%", fig.width=5, fig.height=5}
p1 <- ggcyto(gs.treatment, subset = "P1", aes(y = "FSC-A", x = "FL1-A")) +
    facet_wrap(gRNA ~ cell_line, ncol = 2) +
    geom_hex(bins = 100, aes(fill = ..density..)) +
    geom_vline(xintercept = gfp.min, linetype = "dashed", size = 0.2) +
    # geom_density2d(colour = "gray50", alpha = 0.5, size = 0.4) +
    scale_fill_gradientn(colours = brewer.pal(n = 4, name = "PiYG")) +
    axis_x_inverse_trans() +
    axis_y_inverse_trans() +
    coord_cartesian(ylim = c(-10, 600000), xlim = c(0, 1.4)) +
    scale_y_continuous(labels = scales::scientific_format(digits = 0, prefix = "")) +
    theme_classic() +
    theme(
        panel.background = element_rect(fill = "#f7f7f7"),
        plot.background = element_rect(fill = "transparent", colour = "transparent"),
        aspect.ratio = 1,
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.ticks = element_line(colour = "black", size = LINE.W),
        axis.line = element_line(size = LINE.W),
        axis.title = element_text(size = FONT.SIZE),
        axis.text = element_text(colour = "black", size = FONT.SIZE),
        # axis.text.x=element_blank()
    ) +
    labs(x = "EGFP (FITC-A)", y = "FSC-A", title = "")
p1
ggsave(plot = p1 + guides(fill = "none"), filename = "plots/scatter_EGFP_plt_all.pdf", w = 2.9, h = 2.9, dpi = 300)
```

## Barplot with three replicates
Calculate GFP activation for each condition and merged the stats with assay metadata.

```{r}
sample.names <- gs_pop_get_stats(gs.trans, nodes = c("root"))$sample
sample.stats <- gs_pop_get_stats(
    subset(gs.trans, sample_type == "treatment"),
    type = "percent", nodes = c("GFP")
)
sample.meta <- cyto_details(subset(gs.trans, sample_type == "treatment"))
stat.df <- merge(
    sample.stats, sample.meta,
    by.x = "sample", by.y = "name"
)
stat.df
```

This is also for Supplemenetary Figure.
```{r}
g.barplt <- ggplot(stat.df, aes(y = percent * 100, x = cell_line)) +
    facet_grid(~gRNA) +
    geom_bar(aes(group = gRNA, fill = gRNA_type),
        position = "dodge", stat = "summary", fun = "mean"
    ) +
    geom_point(colour = "black", size = 2.0, shape = 1) +
    theme_classic() +
    scale_x_discrete(expand = c(0, 0), labels = c("gRNA_WT001" = "BC1", "gRNA_WT002" = "BC2")) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 5)) +
    labs(x = "", y = "EGFP positive cells (%)") +
    scale_fill_manual(values = c("gray", "green3")) +
    theme(
        panel.background = element_rect(fill = "transparent", colour = "transparent"),
        plot.background = element_rect(fill = "transparent", colour = "transparent"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        axis.ticks = element_line(colour = "black", size = LINE.W),
        axis.line = element_line(size = LINE.W),
        axis.title = element_text(size = FONT.SIZE),
        axis.text.y = element_text(colour = "black", size = FONT.SIZE),
        axis.text.x = element_text(colour = "black", size = FONT.SIZE, angle = 90, hjust = 1, vjust = 0.5)
    ) +
    guides(fill = "none")
ggsave(plot = g.barplt, filename = "./plots/Indel_mean_EGFP_02042022.pdf", w = 1.0, h = 1.8)
g.barplt
```


