---
title: Error-activation analysis of indel assay
author: "Soh Ishiguro"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: FALSE
    toc_depth: 3
    toc_collapsed: FALSE
    number_sections: yes
    keep_md: yes
    fig_caption: yes
  pdf_document: default
---

```{r knitopt, include=FALSE}
knitr::opts_chunk$set(
    comment = "#>",
    out.width = "50%",
    dpi = 300,
    progress = FALSE
)
```

# Background
This notebook provides error-activation analysis for indel reporter assay (CA058). To load `GatingSet` object, I will use achived data `/Users/mathilda/Dropbox (Yachie Lab)/SOH_ISHIGURO.LAB/Projects/ClonSelect/Flow_data_analysis/CA058_02042022/notebook/gs_achirve`. Live cell gate was defined by the script of `CA058_analysis.rmd`.

```{r libs, message=FALSE}
library(flowWorkspace)
library(CytoExploreR, include.only = "cyto_details")
library(rmarkdown)
library(scales)
library(RColorBrewer)
library(ggplot2)
library(ggridges)
library(ggforce)
library(dplyr)
library(pracma)

setwd("./notebook/")
```

# Load GatingSet object
```{r transform, cache=FALSE}
inv.log.trans <- function(x) {
    ts <- flowWorkspace::flowjo_log_trans()
    inv <- ts[["inverse"]](x)
    return(inv)
}
log.trans <- function(x) {
    ts <- flowWorkspace::flowjo_log_trans()
    trs <- ts[["transform"]](x)
    return(trs)
}

# Load GatingSet
gs.trans <- load_gs("./gs_achieve", verbose = TRUE)
gs_get_pop_paths(gs.trans)
sampleNames(gs.trans)
```

# Generate dynamic gate
```{r dynamic gate generation, cache=FALSE, message=FALSE}
gs.positive <- subset(gs.trans, gRNA_type == "gRNA_OT")
gs.negative <- subset(gs.trans, gRNA_type == "gRNA_NT")

bins <- seq(0, 1e5, by = 50)
bins.trans <- log.trans(bins)

compute_dynamic_gate <- function(gs, inv.bins) {
    df.stat <- data.frame()
    for (min.cutoff in inv.bins) {
        # Create dynamic FL1-A gate
        rg_gfp <- flowCore::rectangleGate(
            "FL1-A" = c(min.cutoff, Inf), "FSC-A" = c(-Inf, Inf),
            filterId = "GFP_dynamic"
        )
        gates <- flowWorkspace::gs_get_pop_paths(gs)
        if ("/P1/GFP_dynamic" %in% gates) {
            flowWorkspace::gs_pop_remove("GFP_dynamic", gs = gs)
            flowWorkspace::gs_pop_add(gs, rg_gfp, parent = "P1")
            flowWorkspace::recompute(gs)
        } else {
            flowWorkspace::gs_pop_add(gs, rg_gfp, parent = "P1")
            flowWorkspace::recompute(gs)
        }
        stat <- flowWorkspace::gs_pop_get_stats(gs, "GFP_dynamic", type = "percent") %>%
            dplyr::mutate(percent, cutoff = inv.log.trans(min.cutoff))
        df.stat <- rbind(df.stat, stat)
    }
    return(df.stat)
}

df.positive <- compute_dynamic_gate(gs.positive, bins.trans[1:100]) # use lower 100 cutoff values
df.negative <- compute_dynamic_gate(gs.negative, bins.trans[1:100])
```

We then merge expected positive and negative data.frame of `df.positive` and `df.negative` to `df.mearged`. 

```{r merge_data, message=FALSE}
sample.meta.pos <- cyto_details(gs.positive)
sample.meta.neg <- cyto_details(gs.negative)
stat.df.pos <- merge(df.positive, sample.meta.pos, by.x = "sample", by.y = "name")

# Group two expected negative cell lines and calculate mean GFP percent
stat.df.neg <- merge(df.negative, sample.meta.neg, by.x = "sample", by.y = "name") %>%
    dplyr::group_by(cutoff, cell_line, gRNA_type, rep) %>%
    dplyr::summarise(mean.percent.neg = mean(percent))

df.merged <- merge(stat.df.pos, stat.df.neg,
    by = c("cutoff", "rep", "cell_line")
)
df.merged %>%
    rename(activation = percent, error = mean.percent.neg) %>%
    arrange(error) -> df.merged

df.merged %>%
    head(n = 20) %>%
    paged_table()
```

```{r fill_extrapolated_value}
# Add (0.0, 0.0) data point
cls <- c("WT001", "WT002")
for (cl in cls) {
    df.merged %>%
        dplyr::add_row(
            rep = c("rep1", "rep2", "rep3"), cell_line = c(rep(cl, 3))
    #        activation = c(0.0, 0.0, 0.0), error = c(0.0, 0.0, 0.0), cutoff = c(0.0, 0.0, 0.0)
        ) %>%
        dplyr::arrange(error) -> df.merged
}
```

```{r plt_AUC, fig.width=3, fig.height=2.3}
FONT.SIZE <- 7
LINE.W <- 0.232 # 0.5pt in Keynote
DATE <- "04062022"

plt.AUC <- function(df, filename, pal) {
    g <- ggplot(df, aes(y = activation, x = error)) +
        # facet_zoom(xlim = c(0, 0.05), ylim = c(0, 0.2), show.area = FALSE) +
        geom_abline(intercept = 0, linetype = "dashed", colour = "gray50", size = LINE.W) +
        geom_line(aes(colour = rep), size = .6) +
        labs(x = "%EGFP error", y = "%EGFP activation") + # x:error, y:activation
        scale_y_continuous(limits = c(0, 0.1), labels = scales::percent_format(suffix = ""), expand = c(0, 0)) +
        scale_x_continuous(limits = c(0, 0.1), labels = scales::percent_format(suffix = ""), expand = c(0, 0)) +
        theme_classic() +
        scale_colour_manual(values = rev(brewer.pal(n = 9, name = pal))) +
        theme(
            panel.spacing.x = unit(0.3, "lines"),
            plot.title = element_text(colour = "black", size = FONT.SIZE),
            strip.background = element_rect(colour = "transparent", fill = "transparent"),
            plot.background = element_rect(colour = "transparent", fill = "transparent"),
            panel.background = element_rect(colour = "transparent", fill = "transparent"),
            panel.border = element_blank(),
            axis.title = element_text(colour = "black", size = FONT.SIZE),
            axis.ticks = element_line(colour = "black", size = LINE.W),
            axis.line = element_line(colour = "black", size = LINE.W),
            axis.text = element_text(colour = "black", size = FONT.SIZE)
        ) +
        guides(colour = "none")
    return(g)
}

WT001.df <- df.merged %>% dplyr::filter(cell_line == "WT001")
WT002.df <- df.merged %>% dplyr::filter(cell_line == "WT002")
g1 <- plt.AUC(WT001.df, "test", "Blues")
g2 <- plt.AUC(WT002.df, "test", "Blues")

g1
g2

ggsave(plot = g1, filename = "WT001_AUC_07182022.pdf", w = 2.0, h = 1.95)
ggsave(plot = g2, filename = "WT002_AUC_07182022.pdf", w = 2.0, h = 1.95)
```

We calculate AUC (mean of replicates) for each target. 
```{r calc_AUC}
df.merged %>%
    dplyr::group_by(cell_line, rep) %>%
    dplyr::summarise(auc = trapz(error, activation)) %>%
    group_by(cell_line) %>%
    summarise(mean.auc = mean(auc)) -> df.auc
df.auc
```

