---
title: scCloneSelect flow analysis - scCSExp24
author: "Soh Ishiguro"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: FALSE
    toc_float: yes
    toc_depth: 3
    toc_collapsed: FALSE
    number_sections: yes
    keep_md: yes
    fig_caption: yes
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(
    comment = "#>",
    out.width = "50%",
    dpi = 300,
    warning = FALSE,
    message = FALSE
)
```

# Overview
Flow assay was performed by RinaS (scCSExp24) with the following experimental condition:   
https://www.dropbox.com/s/srpbn72an3r06x2/Exp_20210804_1_scCSExp24_sample_info.tsv?dl=0

This assay (n=1) was using HEK293T cells, to see barcode orthogonality in scCloneSelect reporters.

# Flow analysis
```{r}
library(RColorBrewer)
library(ggplot2)
library(ggcyto)
library(CytoExploreR)
library(flowWorkspace)
library(dplyr)
library(rmarkdown)
library(scales)
library(lemon)
```

Load samples and create GatingSet object
```{r}
cs <- load_cytoset_from_fcs(path = "./fcs")
gs <- GatingSet(cs)
sampleNames(gs)
markernames(gs)
cs_get_cytoframe(cs, "Tube.1.fcs") # check data
```

Apply transformation for fluorescence channels
```{r}
trans.log <- cyto_transformer_log(gs, channels = c("FL1-A", "FL5-A"), plot = FALSE)
gs.trans <- cyto_transform(gs, trans = trans.log, plot = FALSE)
```

Assign metadata to the `gs.trans`
```{r}
meta.df <- read.csv("./Exp_20210804_1_scCSExp24_sample_info.tsv", header = TRUE, sep = "\t", , skip = 1)
sample.df <- cyto_details(gs.trans)
merged.df <- merge(meta.df, sample.df, by.x = "File", by.y = "name")
cyto_details(gs.trans)$ExpID <- merged.df$ExpID
cyto_details(gs.trans)$Cell_line <- merged.df$Cell_line
cyto_details(gs.trans)$gRNA <- merged.df$gRNA
cyto_details(gs.trans)$Target_type <- merged.df$Target_type
cyto_details(gs.trans)$Sample_type <- merged.df$Samle_type
cyto_details(gs.trans) %>% paged_table()
```

Apply live-cell gate (P1)
```{r, eval=FALSE}
cyto_gate_draw(
    gs.trans,
    parent = "root",
    alias = "P1",
    channels = c("FSC-A", "SSC-A"),
    gatingTemplate = "scCSExp24_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 1 * 10^6),
    type = "polygon"
)
```

Apply single-cell gate (P1/P2)
```{r, eval=FALSE}
cyto_gate_draw(
    gs.trans,
    parent = "P1",
    alias = "P2",
    channels = c("FSC-A", "FSC-Width"),
    gatingTemplate = "scCSExp24_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 6 * 10^3),
    type = "polygon"
)
```
Apply GFP gate (P1/P2/GFP)
```{r, eval=FALSE}
cyto_gate_edit(
    gs.trans["Tube.10.fcs"], # use it for negative control reference
    parent = "P2",
    alias = "GFP",
    channels = c("FL1-A", "FSC-A"),
    gatingTemplate = "scCSExp24_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^7),
    ylim = c(-10, 1 * 10^6),
    type = "interval"
)
```

Apply all gates for each sample
```{r}
gt <- gatingTemplate("scCSExp24_Gate_03152022.csv")
gates <- flowWorkspace::gs_get_pop_paths(gs)
if ("/P1/P2/GFP" %in% gates) {
    gs_pop_remove(gs.trans, "P1") # remove existing gates
}
gs_get_pop_paths(gs.trans)
gt_gating(gt, gs.trans)
cyto_plot_gating_scheme(gs.trans[[1]])
```

Merge %GFP stats and sample metadata
```{r}
GFP.stat <- gs_pop_get_stats(gs.trans, node = "GFP", type = "percent") %>% mutate(percent = percent * 100)
df.stat <- merge(cyto_details(gs.trans), GFP.stat, by.x = "name", by.y = "sample")
df.stat <- df.stat %>% mutate(
    col_var = paste(Cell_line, gRNA, sep = ":"),
    gRNA = factor(gRNA, levels = c("BC2gRNA", "BC4gRNA", "BC6gRNA", "BC8gRNA", "BC10gRNA", "BC12gRNA")),
    Cell_line = factor(Cell_line, levels = c("BC2", "BC4", "BC6", "BC8", "BC10", "BC12"))
)
df.stat %>% paged_table()
write.csv(df.stat, quote = FALSE, row.names = FALSE, file = "pop_stat_03152022.csv")
```

# Visualization
```{r}
# define const for visualization
FONT.SIZE <- 9
LABEL.FONT.SIZE <- 8
LINE.W <- 0.3481 # equivalent to 0.75pt in Keynote
```

```{r}
g.heat <- ggplot(
    df.stat %>% dplyr::filter(gRNA != "NA", gRNA != "NTgRNA"),
    aes(y = Cell_line, x = gRNA)
) +
    geom_tile(aes(fill = percent)) +
    scale_fill_gradientn(limits = c(0, 60), colours = c("black", "green3", "green2")) +
    theme_classic() +
    theme(
        axis.text.x = element_text(angle = 90, vjust = 0.5),
        aspect.ratio = 1
    )
g.heat
```

```{r, fig.width=2.3, fig.height=3}
g1 <- ggplot(df.stat %>% dplyr::filter(Target_type != "Control"), aes(y = percent, x = Cell_line)) +
    facet_rep_wrap(~gRNA, ncol = 1, repeat.tick.labels = "left") +
    geom_bar(aes(fill = Target_type), stat = "identity", width = 0.8) +
    theme_classic() +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 50), breaks = c(0, 25, 50)) +
    scale_x_discrete(expand = c(0, 0)) +
    scale_fill_manual(values = c("Match" = "green3", "Mismatch" = "gray")) +
    labs(x = "", y = "EGFP positive cells (%)") +
    theme(
        panel.spacing = unit(-0.3, "lines"),
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        strip.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", size = LABEL.FONT.SIZE),
        axis.text.y = element_text(colour = "black", size = LABEL.FONT.SIZE),
        axis.title = element_text(size = FONT.SIZE),
        axis.line = element_line(size = LINE.W, colour = "black"),
        axis.ticks = element_line(size = LINE.W, colour = "black"),
        # axis.ticks.x = element_blank()
    ) +
    guides(fill = "none")
g1
ggsave(plot = g1, filename = "GFP_pop_BC-gRNA_03152022.pdf", w = 2.3, h = 3)
```

