---
title: Flow analysis of scCloneSelect (Exp_20210719_1)
subtitle: original vs 3' capture reporter comparison
author: "Soh Ishiguro"
date: "`r Sys.Date()`"
output:
  html_document: 
    toc: yes
    toc_float: FALSE
    toc_depth: 3
    toc_collapsed: FALSE
    number_sections: yes
    keep_md: yes
    fig_caption: yes
  pdf_document: default
---

```{r, include=FALSE}
knitr::opts_chunk$set(
    comment = "#>",
    out.width = "50%",
    dpi = 300,
    warning = FALSE,
    message = FALSE
)
```

# Overview
Flow assay was performed by RinaS (Exp_20210719_1) with the following experimental condition:   
`SOH_ISHIGURO.LAB/Projects/ClonSelect/Flow_data_analysis/mESC_flow_Rina/Exp_20210719_1/Exp_20210719_1_sample_info.tsv`

Assay was peformed with n=3, to see GFP activation in scCloneSelect BC2 reporter and original reporter.

# Analysis
```{r}
library(ggplot2)
library(RColorBrewer)
library(ggcyto)
library(CytoExploreR)
library(flowWorkspace)
library(dplyr)
library(rmarkdown)
library(scales)
library(ggridges)
```

Load samples and create GatingSet object
```{r}
cs <- load_cytoset_from_fcs(path = "./fcs")
gs <- GatingSet(cs)
sampleNames(gs)
markernames(gs)
cs_get_cytoframe(cs, "Tube.1.fcs") # check data
```

Apply transformation for fluorescence channels
```{r}
trans.log <- cyto_transformer_log(gs, channels = c("FL1-A", "FL5-A"), plot = FALSE)
gs.trans <- cyto_transform(gs, trans = trans.log, plot = FALSE)
```

Assign metadata to the `gs.trans`
```{r}
meta.df <- read.csv("./Exp_20210719_1_sample_info.tsv", header = TRUE, sep = "\t", , skip = 1)
sample.df <- cyto_details(gs.trans)
merged.df <- merge(meta.df, sample.df, by.x = "File", by.y = "name")
cyto_details(gs.trans)$Cell_line <- merged.df$Cell_line
cyto_details(gs.trans)$gRNA <- merged.df$gRNA
cyto_details(gs.trans)$Rep <- merged.df$Rep
cyto_details(gs.trans)$Sample_type <- merged.df$Sample_type
cyto_details(gs.trans)$Target_type <- merged.df$Target_type
cyto_details(gs.trans)$File <- merged.df$File
cyto_details(gs.trans)$ExpID <- merged.df$ExpID
```

Check all metatada and fcs files assigned.
```{r}
cyto_details(gs.trans) %>%
    arrange(name) %>%
    paged_table()

gs_pop_get_stats(gs.trans) %>% paged_table()
```

Create manual gate
```{r, eval=FALSE}
# Live cell gate
cyto_gate_draw(
    gs.trans,
    parent = "root",
    alias = "P1",
    channels = c("FSC-A", "SSC-A"),
    gatingTemplate = "scCS_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 1 * 10^6),
    type = "polygon"
)
```

```{r, eval=FALSE}
# Single-cell gate
cyto_gate_draw(
    gs.trans,
    parent = "P1",
    alias = "P2",
    channels = c("FSC-A", "FSC-Width"),
    gatingTemplate = "scCS_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^6),
    ylim = c(-10, 6 * 10^3),
    type = "polygon"
)
```

```{r, eval=FALSE}
# GFP gate
cyto_gate_draw(
    gs.trans["Tube.7.fcs"], # use it for negative control reference
    parent = "P2",
    alias = "GFP",
    channels = c("FL1-A", "FSC-A"),
    gatingTemplate = "scCS_Gate_03152022.csv",
    xlim = c(-10, 1 * 10^7),
    ylim = c(-10, 1 * 10^6),
    type = "interval"
)
```

Apply all gates for each sample
```{r}
gt <- gatingTemplate("scCS_Gate_03152022.csv")
gates <- flowWorkspace::gs_get_pop_paths(gs)
if ("/P1/P2/GFP" %in% gates) {
    gs_pop_remove(gs.trans, "P1")
}
gt_gating(gt, gs.trans)
plot(gt)
cyto_plot_gating_scheme(gs.trans[[1]])
```

Merge metadata and %GFP
```{r}
GFP.stat <- gs_pop_get_stats(gs.trans, node = "GFP", type = "percent") %>% mutate(percent = percent * 100)
df.stat <- merge(cyto_details(gs.trans), GFP.stat, by.x = "File", by.y = "sample")
df.stat <- df.stat %>% mutate(col_var = paste(Cell_line, gRNA, sep = ":"))

# Assign custom colours
df.stat %>% mutate(col = case_when(
    col_var == "BC2_GTG:BC2_gRNA" ~ "green3",
    col_var == "BC2_ATG:BC2_gRNA" ~ "green3", # dodgerblue2
    col_var == "scCS_BC2_GTG:BC2_gRNA" ~ "green3",
    col_var == "scCS_BC2_ATG:BC2_gRNA" ~ "green3", # dodgerblue2
    col_var == "BC2_GTG:NT_gRNA" ~ "gray",
    col_var == "BC2_ATG:NT_gRNA" ~ "gray",
    col_var == "scCS_BC2_GTG:NT_gRNA" ~ "gray",
    col_var == "scCS_BC2_ATG:NT_gRNA" ~ "gray",
)) -> df.stat

df.stat %>% paged_table()
write.csv(df.stat, quote = FALSE, row.names = FALSE, file = "pop_stat_03152022.csv")
```

Calculate MFI for reporter cell lines
```{r}
df.MFI <- cyto_stats_compute(gs.trans, channels = c("FL1-A"), alias = "GFP", stat = "median") %>% dplyr::filter(Sample_type == "Treatment")
```

# Visualization
```{r}
# define const for visualization
FONT.SIZE <- 9
LABEL.FONT.SIZE <- 8
LINE.W <- 0.3481 # equivalent to 0.75pt in Keynote
```

```{r, fig.width=1.2, fig.height=2.7, out.width="20%", fig.show="hold"}
plt.bar <- function(df, ...) {
    arglist <- list(...) # additional ggplot2 object(s)
    g <- ggplot(df) +
        facet_wrap(~Cell_line) +
        geom_bar(aes(group = gRNA, fill = col, x = gRNA, y = percent), position = "dodge", stat = "summary", fun = "mean") +
        geom_point(aes(y = percent, x = gRNA), colour = "black", size = 2.0, shape = 1, stat = "identity") +
        # scale_fill_manual(values = c("NT_gRNA" = "gray", "BC2_gRNA" = "green3")) +
        scale_fill_identity() +
        scale_x_discrete(expand = c(0, 0), labels = c("OT gRNA", "NT gRNA")) +
        scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
        theme_classic() +
        theme(
            strip.background = element_blank(),
            strip.text.x = element_blank(),
            strip.text.y = element_blank(),
            axis.text.x = element_text(colour = "black", angle = 90, hjust = 1, vjust = 0.5, size = LABEL.FONT.SIZE),
            axis.text.y = element_text(colour = "black", size = LABEL.FONT.SIZE),
            axis.title = element_text(size = FONT.SIZE),
            axis.line = element_line(size = LINE.W, colour = "black"),
            axis.ticks = element_line(size = LINE.W, colour = "black"),
        ) +
        labs(y = "EGFP positive cells (%)", x = "") +
        guides(fill = "none")

    # overwrite existing layers
    if (length(arglist) > 0) {
        return(g + arglist)
    } else {
        return(g)
    }
}

# reporter cells
df.GTG <- df.stat %>% dplyr::filter(Sample_type == "Treatment")
g.bar.GTG <- plt.bar(df.GTG)

# for ATG ctrl cells
df.ATG <- df.stat %>% dplyr::filter(Sample_type == "Control")
g.bar.ATG <- plt.bar(df.ATG, ggplot2::scale_y_continuous(expand = c(0, 0), limits = c(0, 99)))

w <- 1.2
h <- 2.7
ggsave(plot = g.bar.GTG, filename = "./plots/GFP_barplt_GTG_cells.pdf", w = w, h = h)
ggsave(plot = g.bar.ATG, filename = "./plots/GFP_barplt_ATG_cells.pdf", w = w, h = h)
plot(g.bar.GTG + labs(subtitle = "GTG cells")) # left: original, right: scCloneSelect
plot(g.bar.ATG + labs(subtitle = "ATG ctrl cells")) # left: original, right: scCloneSelect
```

```{r, fig.height=2, fig.width=1, out.width="20%"}
g.MFI <- ggplot(df.MFI %>% dplyr::filter(gRNA == "BC2_gRNA"), aes(y = MedFI, x = Cell_line)) +
    geom_bar(aes(group = gRNA), fill = "orange", stat = "summary", fun = "mean", position = "dodge", width = 0.7) +
    geom_point(colour = "black", size = 2, shape = 1) +
    theme_classic() +
    scale_x_discrete(expand = c(0, 0), label = c("BC2_GTG" = "Original\nreporter", "scCS_BC2_GTG" = "3' capture\nreporter")) +
    scale_y_continuous(
        trans = "log10",
        labels = trans_format("log10", math_format(10^.x)),
        breaks = trans_breaks("log10", function(x) 10^x),
        expand = c(0, 0),
        limits = c(1, 1e6)
    ) +
    labs(x = "", y = "Median EGFP intensity (a.u.)") +
    theme(
        strip.background = element_blank(),
        strip.text.x = element_blank(),
        strip.text.y = element_blank(),
        axis.text.x = element_text(colour = "black", angle = 90, hjust = 1, vjust = 0.5, size = 7),
        axis.text.y = element_text(colour = "black", size = 7),
        axis.title = element_text(size = FONT.SIZE),
        axis.line = element_line(size = LINE.W, colour = "black"),
        axis.ticks = element_line(size = LINE.W, colour = "black"),
    )
ggsave(plot = g.MFI, filename = "./plots/GFP_MFI_03182022.pdf", h = 2.12, w = 0.79)
```
